---
title: "A tribute to Serpil Hamdi Tüzün"
author: "by Orhan Aktas"
date: '2022-08-04'
output:   
    html_document:
      df_print: !expr knitr::kable
      toc: yes
      number_sections: yes
      toc_depth: 2
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  class.source = "bg-warning text-warning"
)
```

#### Required Libraries {.unlisted .unnumbered}

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(rvest)
library(ggthemes)
library(ggtext)
library(tidyquant)

```

# Introduction

# get year by year squad data from transfermarkt

function definition
```{r}
my_function <- function(year) {
  
  preurl<-"https://www.transfermarkt.com.tr/be%C5%9Fikta%C5%9F-jk/kader/verein/114/plus/1/galerie/0?saison_id="
  url <- paste0(preurl,as.character(year))
  scraped_url <- read_html(url)
  
  PlayerNames_bjk  <- scraped_url %>% 
    html_nodes(".hauptlink a") %>% 
    html_text() %>% 
    as_tibble() %>% 
    rename(names=value) %>% 
    mutate(names=str_trim(names))
  
  Playerages_bjk  <- scraped_url %>% 
    html_nodes("#yw1 td:nth-child(3)") %>% 
    html_text() %>% 
    as_tibble() %>% 
    rename(ages=value) 
  
  Playerdates_bjk  <- scraped_url %>% 
    html_nodes("td:nth-child(8)") %>% 
    html_text() %>% 
    as_tibble()%>% 
    rename(joined_date=value)
  
  Playerpreteams_bjk  <- scraped_url %>% 
    html_nodes("td:nth-child(9)") %>% html_element("a") %>% 
    html_attr("title") %>% 
    as_tibble() %>% 
    mutate(pre_team=str_sub(value, 1, str_locate(value,": Ablöse") [,"start"]-1)) %>% 
    select(-value)
  
  tibble(PlayerNames_bjk,Playerages_bjk,Playerdates_bjk,Playerpreteams_bjk) 
}
```

define years
```{r}
years <- 1969:2021
```


scrape the data and write to a local file
```{r eval=FALSE}
squads <- years %>%
  map(my_function) %>%
  set_names(years)

write_rds(squads,"data/squads.rds")
rm(squads)
```

read squad data from local file
```{r}
squads<-read_rds("data/squads.rds")
```

add age joined to the club
```{r}

add_age_joined <- function(squad) {
  squad %>% 
    mutate(ages=str_replace(ages, "\\s*\\([^\\)]+\\)", "")) %>% 
    mutate(year_born=as.integer(str_sub(ages, -4)),
         year_joined=as.integer(str_sub(joined_date, -4)),
         age_joined=year_joined-year_born) %>%  
    select(age_joined)
}

squads2<-squads %>% map(add_age_joined)

```


visualize ozkaynak based 
```{r}
ozk_oyuncu<- squads %>% 
  map(mutate,Ozkaynak=ifelse(grepl('Beşiktaş', pre_team)==TRUE,1,0)) %>% 
  map(summarise,sum(Ozkaynak)) %>% 
  bind_rows() %>% 
  bind_cols(years %>% as_tibble) %>% 
  set_names("Ozkaynak","year") %>% 
  relocate(year) 

ozk_oyuncu %>%  ggplot(aes(x=year))+
                  geom_col(aes(y=Ozkaynak))
```

visualize number of players who joined to the club before age of 22
```{r}
oyuncu_22<- squads2 %>% 
  map(mutate,youth=ifelse(age_joined<22,1,0)) %>% 
  map(summarise,sum(youth,na.rm = TRUE)) %>% 
  bind_rows() %>% 
  bind_cols(years %>% as_tibble) %>% 
  set_names("youth","year") %>% 
  relocate(year) 

oyuncu_22 %>%  ggplot(aes(x=year))+
                  geom_col(aes(y=youth))
```

# get year by year season data from wikipedia and join with the previous
```{r eval=FALSE}
wiki_url<-"https://tr.wikipedia.org/wiki/Be%C5%9Fikta%C5%9F_(futbol_tak%C4%B1m%C4%B1)_sezonlar%C4%B1_listesi"
scraped_wiki_url <- read_html(wiki_url)

scraped_wiki_url %>%
  html_nodes(".wikitable:nth-child(9) td:nth-child(9) , .wikitable:nth-child(9) td:nth-child(8), .wikitable:nth-child(9) td:nth-child(7), .wikitable:nth-child(9) td:nth-child(6), .wikitable:nth-child(9) td:nth-child(5), .wikitable:nth-child(9) td:nth-child(4), .wikitable:nth-child(9) td:nth-child(3), :nth-child(9) th:nth-child(1) a") %>%
  html_text() %>%
  as_tibble() %>%
  bind_cols(rep(c("year", "O", "G", "B", "M", "A", "Y", "PUAN"), 62)) %>%
  bind_cols(ceiling(1:(8 * 62) / 8)) %>%
  set_names("value", "var", "id") %>%
  mutate(value = str_trim(value)) %>%
  pivot_wider(names_from = var, values_from = value) %>%
  slice(6:n()) %>%
  rename(sezon = year, year = id) %>%
  mutate(year = as.integer(year + 1957))%>%
  rows_update(tibble(
    year = 2011, sezon = "2011-12", O = "40", G = "16", B = "12", M = "12", A = "55", Y = "47", PUAN = "60"
  ),
   by="year"
  ) %>% 
  mutate_at(vars(O:PUAN), as.integer) %>% 
  mutate(PuanOrt=(G*3+B)/O) %>% 
  mutate(tuzun=ifelse(year>1974&year<1990,1,0),
         seba=ifelse(year>1983&year<2000,1,0),
         gordon=ifelse(year>1986&year<1994,1,0)) %>% 
  left_join(ozk_oyuncu,by="year") %>% 
  left_join(oyuncu_22,by="year") %>% 
  write_rds("data/main_table.rds")


```

read squad data from local file
```{r}
main_table<-read_rds("data/main_table.rds")
```


visualize
```{r}
main_table %>% 
  filter(year>1970&year<2000)%>%
  mutate(GoalOrt=A/O) %>% 
  ggplot(aes(x=year))+
    geom_col(aes(y=Ozkaynak),fill="black")+
    geom_col(aes(y=youth),fill="black",alpha=0.5)+
    geom_line(aes(y=(PuanOrt-1)*10),col="red",size=2)+
    # serpil hamdi tuzun era
    geom_errorbarh(
      data=main_table %>% filter(tuzun==1),
      aes(xmax = max(year), xmin = min(year), y=1, height = .2),
      color="red")+
    geom_label(data=main_table %>% filter(tuzun==1),
               aes(label=str_wrap("Serpil Hamdi Tüzün",20), x=mean(year),y=1), 
               size=3,
               alpha=0.2) +
    # suleyman seba era
    geom_errorbarh(
      data=main_table %>% filter(seba==1),
      aes(xmax = max(year), xmin = min(year), y=2, height = .2),
      color="red")+
    geom_label(data=main_table %>% filter(seba==1),
             aes(label=str_wrap("Süleyman Seba",20), x=mean(year),y=2), 
             size=3,
             alpha=0.2) +
    # gordon milne era
    geom_errorbarh(
      data=main_table %>% filter(gordon==1),
      aes(xmax = max(year), xmin = min(year), y=3, height = .2),
      color="red")+
    geom_label(data=main_table %>% filter(gordon==1),
           aes(label=str_wrap("Gordon Milne",20), x=mean(year),y=3), 
           size=3,
           alpha=0.2) +
  theme_minimal()+
  # 
    scale_x_continuous(breaks=seq(1971,1999,1),labels=as.character(71:99)) 

    #theme(axis.text.x =element_text( angle=90))
  
```

```{r}
positions_by_year <- read_excel("C:/Users/user/Desktop/positions_by_year.xlsx")

positions_by_year %>% 
  ggplot(aes(x=year,y=position))+
    geom_line(size=2,alpha=0.7)+
    geom_ma(ma_fun = SMA, n = 3) + 
    theme_minimal() +
    scale_y_continuous(breaks=seq(11, 1, -1),trans = "reverse") +   

  theme(legend.position="none")+
  theme(
          panel.background = element_blank(),    
          axis.text.y = element_text(angle = 0, vjust = -.5, 
                                     margin = margin(r = -30)),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_line(colour = 'light grey', 
                                          linetype = 'dashed',size = .35),
          axis.ticks.x = element_line(color = "light grey"),
          axis.ticks.y = element_blank(),
          axis.line.x = element_line(colour = "light grey", linetype = "solid",
                                     size = .5))+
  # annotations: 
  
  ## Serpil Hamdi Tüzün
  annotate("richtext", 
           label = "Serpil Hamdi<br>Tüzün",
           x = 1967, y = 10, size = 5, color = "black",
           label.colour = NA, fill = NA) +
  annotate(geom = "point", x = 1975, y = 11, colour = "black", size = 3)+
  annotate("curve",
           x = 1975, xend = 1969, 
           y = 11, yend = 10,
           color = "black", size = 1, 
    curvature = -.3, arrow = arrow(length = unit(2, "mm")))+
  
  
  ## Süleyman Seba
  annotate("richtext", 
           label = "Süleyman<br>Seba",
           x = 1985, y = 8, size = 5, color = "black",
           label.colour = NA, fill = NA) +
  annotate(geom = "point", x = 1984, y = 2, colour = "black", size = 3)+
  annotate("curve",
           x = 1984, xend = 1985, 
           y = 2, yend = 7.5,
           color = "black", size = 1, 
    curvature = -.2, arrow = arrow(length = unit(2, "mm")))+
  
  ## Gordon Milne
  annotate("richtext", 
           label = "Gordon<br>Milne",
           x = 1990, y = 6, size = 5, color = "black",
           label.colour = NA, fill = NA) +
  annotate(geom = "point", x = 1987, y = 2, colour = "black", size = 3)+
  annotate("curve",
           x = 1987, xend = 1990, 
           y = 2, yend = 5.5,
           color = "black", size = 1, 
    curvature = -.2, arrow = arrow(length = unit(2, "mm")))




```
same values with a new chart similar to the transfmarkt's


```{r}
# with serpil hamdi tüzün era
positions_by_year %>% 
  ggplot(aes(x=year,y=position))+
    #geom_rect(aes(xmin = 1975,xmax = 1990,ymin = - Inf,ymax = Inf,  fill = "red"),alpha=0.1) +
    #geom_line(size=2,alpha=0.7)+
  #   geom_ma(ma_fun = SMA, n = 3,col="red") + 
  # geom_ma(ma_fun = SMA, n = 5,col="blue") +
  geom_ma(ma_fun = SMA, n = 7,col="green") +
    theme_minimal() +
    scale_y_continuous(breaks=seq(11, 1, -1),trans = "reverse") +   

  theme(legend.position="none")+
  theme(
          panel.background = element_blank(),    
          axis.text.y = element_text(angle = 0, vjust = -.5, 
                                     margin = margin(r = -30)),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_line(colour = 'light grey', 
                                          linetype = 'dashed',size = .35),
          axis.ticks.x = element_line(color = "light grey"),
          axis.ticks.y = element_blank(),
          axis.line.x = element_line(colour = "light grey", linetype = "solid",
                                     size = .5))
```
u-19 milli takım
```{r}

```

